TurtleBot SLAM

### 2013-01-23 trunk/Article/Article.tex
minor changes

\subsubsection{Stage and RViz}
During the development process of a system running on a robot one crucial part is testing. A ROS stack called Stage enables developers to test code on a simulated robot allowing it to move around in virtual environments. Stage visualizes the robot, its movement and measurements. The user can load individually created maps into Stage and specify the robot's configuration. A second important stack needed especially for validation is RViz. It visualizes all requested data. In Figure \ref{fig:stage_and_rviz} Stage and RViz are presented next to each other. While Stage (Left) shows the real map, RViz (Right) presents the map data as seen by the robot.

\subsection{TurtleBot}
\label{subsec:turtle}
The product of this project is designed to run on a mobile robot called TurtleBot (Figure \ref{fig:turtlebot}) \citep{turtlebot}. It is a low-cost, customizable, personal robot kit. The base of the TurtleBot is an iRobot Create (A). It holds a battery pack and a 150 degrees/second Single Axis Gyro. The model shown in Figure \ref{fig:turtlebot} uses a Microsoft Kinect sensor (B), the one used in this paper uses a laser range sensor and a laptop (C) to run all the processes. The robot can also be customized using the mounting hardware (D). The robot runs on an open-source SDK based on ROS which integrates all the software needed to get the TurtleBot running. Some capabilities are included in ROS and can be used in the TurtleBot after minor configurations. These include mapping and navigation, for example. 
\begin{figure}[h]
	\centering
		\includegraphics[width=0.50\textwidth]{figures/turtlebot.png}
	\caption{A TurtleBot}
	\label{fig:turtlebot}
\end{figure}

\subsection{Exploration}
\label{subsec:sotaExplore}
Usually robots localize themselves and navigate through environments using a map which is provided by an external source. In this case navigating from one location to another is an easy problem. If the robot does not have access to a map, navigating through an environment is problematic. The problem treated in this paper (Subsection \ref{sec:problem}) assumes an unknown environment and, hence, no map. The robot needs to be able to explore its environment without any knowledge about it. It needs to rely on the incoming laser and odometry data only. In addition to the creation of the map and the localization of itself (the SLAM problem) the robot needs to explore the environment as efficient as possible. We define a good exploration strategy as one that generates an accurate map in the shortest amount of time.

Ideally the robot updates its internal map of the environment constantly while moving. One idea to efficiently explore the world is to move to locations which maximize the robot's information gain. This target location needs to be extracted from the limited information the robot has collected with its sensors.
To achieve this the robot makes use of the idea of frontier-based exploration, first introduced by Yamauchi \cite{Yamauchi} in 1998. Frontiers are defined to be the boundaries between free and unexplored space. Intuitively, when a robot moves to a frontier it is likely that it makes measurements in previously unexplored space. Given that new data can be added to its map. The explored region grows once the robot reaches a frontier. Moreover an extended map results in new frontiers. By constantly moving to frontiers the robot expands its knowledge about the environment.

@@

Other approaches estimate the complete trajectory from the set of measurements. These approaches address the so called full SLAM problem and they typically rely on least-square error minimization techniques \cite{Leastsquares}.

An intuitive way to formulate the SLAM problem is to consider it as a pose graph optimization problem. The nodes of the graph represent robot poses or landmarks. The edges encode sensor measurements that constraint the connected poses. Once such a graph is constructed the graph-based SLAM, or Graph SLAM, method tries to find a configuration of the nodes that is maximally consistent with the constraints contained in the edges.\\

SLAM is a problem because localization requires a map and mapping requires the location of the device or robot that wants to create a map. There are quite a few approaches that successfully solve the SLAM problem. 
In general there are two interpretations it. The first estimates the current robot pose along with the map. The core of such a system is a filtering technique. New measurements augment and refine the estimate as they become available. Due to their incremental nature these filtering approaches are commonly referred to as on-line SLAM methods. FastSlam \citep{Montemerlo02} is an example of an on-line SLAM implementation. It uses a Particle Filter to approximate the robots pose and the landmark positions on the map. Other methods have the goal to find the trajectory of the robot and the map. These solve the full SLAM problem and typically rely on least-square error minimization techniques \cite{Leastsquares} to find adequate estimates.  
An example of such a an approach is graph-based SLAM or Graph SLAM.

\subsubsection{Graph SLAM}
An intuitive way to formulate the SLAM problem is to view it as a pose graph optimization problem. Each pose along the trajectory of the robot is represented by one node. Edges are labeled with relative spatial information between the connecting nodes. This graph is built incrementally. While the robot is in motion, new nodes are added to the graph. Building up the graph, as explained above, is a straight forward procedure. The essence of the Graph SLAM method, however, is the optimization step. In this step the algorithm tries to find a spatial configuration of the trajectory which is maximally consistent with a system of constraints obtained from the edges. This problem is commonly solved using least-square error minimization techniques \citep{Grisetti}.

@@

The implemented packages are making use of the external library ``Eigen" \citep{eigen}. It provides extensive linear algebra operations. Included are simple matrix operations as well as advanced techniques. Two techniques which are used in this project are Singular Value Decomposition and sparse Cholesky factorization.

\subsection{Graph SLAM}
\label{sec:implSLAM}
The implemented Graph SLAM algorithm is based on \citep{Grisetti}. The graph consists of nodes representing poses along the robot trajectory. Odometry and sensor measurements are incorporated into the constraints contained in the edges. Similar to the Graph SLAM procedure explained in Section \ref{subsec:sotaSlam}, an edge connects two adjacent poses of the robot. 
\begin{figure}[h]
	\centering
		\includegraphics[width=0.50\textwidth]{figures/graph.png}
	\caption{An example of a simple graph \citep{Grisetti}}
	\label{fig:turtlebotexample_graph}
\end{figure}
When the graph is created the pose graph is optimized according to the constraints. This ``optimized" trajectory then determines the positions of obstacles on the map. The laser scan data are refined according to the poses they were obtained at. Each edge, thus, is labeled with the spatial difference between two poses \(x_i\) and \(x_j\). Let this error be denoted by \(e_{ij}(x_i, x_j)\). This information is obtained from a combination of odometry data and laser range sensor data. In fact, the raw odometry data is corrected using a laser scan matcher (see Subsection \ref{sec:scan}). 
Figure \ref{fig:turtlebotexample_graph} shows a sketch of a graph created by the algorithm. Two poses \(x_i\) and \(x_j\) are connected by an edge (Solid arrow). The ellipse at the top illustrates the label of the edge.

This graph-based SLAM approach calculates the most likely path as follows:
First it finds the Jacobians \(A\) and \(B\) of every two estimated poses connected by an edge.
\begin{equation}
\label{eq:jacobi_a}
	A_{ij} = \frac{\partial e_{ij}(x)}{\partial x_i}
\end{equation}
\begin{equation}
\label{eq:jacobi_b}
	B_{ij} = \frac{\partial e_{ij}(x)}{\partial x_j}
\end{equation}
Note that in Equations \ref{eq:jacobi_a} and \ref{eq:jacobi_b} the x represents the current guess of the concerning robot pose. As explained above, \(e_{ij}(x_i, x_j)\) is the error between the two poses \(x_i\) and \(x_j\). The Jacobians \(A\) and \(B\) are then added to a matrix \(H\) (Equation \ref{eq:H}).
\begin{equation}
\label{eq:H}
\begin{aligned}
	H_{[ii]} += A^{T}_{ij} \Omega_{ij} A_{ij} \\
	H_{[ij]} += A^{T}_{ij} \Omega_{ij} B_{ij} \\
	H_{[ji]} += B^{T}_{ij} \Omega_{ij} A_{ij} \\
	H_{[jj]} += B^{T}_{ij} \Omega_{ij} B_{ij} \\
\end{aligned}
\end{equation}
In this equation $\Omega_{ij}$ is the information matrix of a measurement between \(x_i\) and \(x_j\). It is defined as shown in Equation \ref{eq:omega}.
Sigma is the $3 \times 3$ covariance matrix of the measurements. In our implementation the diagonal of $\Sigma$ is filled with arbitrary values.
\begin{equation}
\label{eq:omega}
\Omega = \Sigma^{-1}
\end{equation}
In addition to the matrix $H$ a coefficient vector \(b\) is kept up to date as shown in Equation \ref{eq:b}. This vector together with the \(H\) form a set of linear constraints.
\begin{equation}
\label{eq:b}
\begin{aligned}
	b_{[i]} += A^{T}_{ij} \Omega_{ij} e_{ij} \\
	b_{[j]} += B^{T}_{ij} \Omega_{ij} e_{ij} 
\end{aligned}
\end{equation}
The optimization of the vector \(x = (x_1, ..., x_n)^T \) according to the constraints specified in \(H\) and \(b\) is the core of this approach. This step is performed using sparse Cholesky factorization.

A new node is added to the graph once the robot moves 0.5 meters. The algorithm optimizes the trajectory like explained once a new node is added.\\
In one special case the algorithm adds an extra edge, connecting to non-adjacent nodes, to the graph. This happens when the distance between the two connected robot poses is smaller than a margin. In fact, this special case enables the algorithm to create loops in the graph and more accurately estimate the robots trajectory and the map. This procedure is referred to as ``loop closing" (see Subsection \ref{sec:loopClosing}). 
\subsection{Scan matching}
\label{sec:scan}
The scan matcher, used by this program, is necessary to acquire an optimized movement vector between two robot poses. This is needed by the Graph SLAM algorithm (see Subsection \ref{sec:implSLAM}). The scan-matcher realized for this project can be divided into two steps. At the beginning, the two scans are matched making use of raw odometry data, retrieved by the robot's motion sensor. This is improved upon by running an iteration of the iterative closest point algorithm (ICP) \citep{scanmatch}. It tries to align the two scans minimizing the total distance of each point in the first scan to its nearest neighbor in the second.

The general procedure of the algorithm \citep{scanmatch} is to first find the nearest neighboring point in the second scan for all point in the first. To achieve this nearest neighbor search is used.
In this implementation multiple points can have the same corresponding neighbor-point. If there is no point from the other scan within a specified radius of one point, it is excluded. These outliers would influence further calculations negatively.
The next step in the process is to calculate the centroids of both clouds. These are subtracted from every point in the concerning cloud. In that way the clouds are centered at the origin. The calculated points are then added to two separate $2 \times n$ matrices, where $n$ is the number of neighbor pairs. Next the rotation matrix which transforms one cloud to the other is needed. The Singular Value Decomposition (SVD) of the product of the two previously mentioned matrices returns three new matrices $U$, $\Sigma$ and $V$. The required rotation matrix R can be obtained as described in Equation \ref{eq:RVUT}.
\begin{equation}
\label{eq:RVUT}
 R = V * U^T
\end{equation}
The translation vector $t$ is the difference between the two centroids. $R$ and $t$ together form a rigid transformation. This transformation matches the two scans. Moreover it is used to correct the robot's odometry data.

@@

\subsection{Mapping}
\label{sec:mapping}
The map is internally represented as an occupancy grid. It consists of cells which hold information about obstacles, free/- and unknown space.

The occupancy grid is a structure that is represented by a one-dimensional array. The size of an occupancy grid has to be tuned problem specific. A too large grid can lead to performance problems. If the grid is too small, however, the map will be inaccurate. Fitting a map on a smaller grid results in a low resolution.
Internally the program represents the grid as a two dimensional array. This makes navigating to either side more intuitive. Each grid cell contains an integer value. The value-range for ROS' occupancy grid lies between $-1$ and $100$ \cite{occupancy}. The values indicate the state as follows:
\begin{description}
\item{$-1$} : Unknown space, not yet covered by robot's perception sensors
\item{$0$} : Known free space, robot can drive on it safely
\item{$1-100$} : Probability of being an obstacle, $1$ being a low probability and $100$ being a definite obstacle.
\end{description}
\begin{figure}[htbp]
	\centering
		\includegraphics[width=0.45\textwidth]{figures/Occup.png}
	\caption{A visualized occupancy grid}
	\label{fig:Occupancy}
\end{figure}
Figure \ref{fig:Occupancy} shows an example occupancy grid. The robot stands in the center of it and makes one measurement consisting of several laser scans. Each scan returns the range it was able to reach. The returned range is either the maximally possible range of the laser or a lower value, indicating an obstacle (value of 100). The range is transformed into global coordinates with respect to the robot's location, orientation and the bearing of the laser beam. The resulting point is then transferred into the occupancy grid as an obstacle if the maximum range was not reached. The cells lying in a straight line between the transformed range and the robot are considered free space (value of 0), since the laser was able to pass through this area. All the cells that are not affected by any of the previous steps remain unexplored (value of -1). Naturally, this is the case for cells behind obstacles, out of range and directly behind the robot.

@@

\subsection{Exploration}
\label{sec:wfd}
One part of the assignment is autonomous exploration. The robot has to drive autonomously through an unknown world. This feature is implemented using Wavefront frontier detection (WFD).

WFD is a frontier-based exploration algorithm \citep{Keidar}. It uses four different data structures. These store map and frontier information during the search process. They are:
\begin{enumerate}
\item{Map-Open-List}
\item{Map-Close-List}
\item{Frontier-Open-List}
\item{Frontier-Close-List}
\end{enumerate}
To understand the need of these four structures it is necessary to know that WFD is using a breadth first search (BFS), starting at the current robot position. Moreover frontier detection requires the robot to have an internal representation of the current environment. In this implementation an occupancy grid (Subsection \ref{sec:mapping}) is used.
Upon launching the algorithm it first searches for unoccupied space around the current robot location. These grid cells are saved in the Map-Open-List. 
Frontiers are found by making use of the occupancy grid's structure. The criterion for a map-point to be a frontier is the existence of both, unknown space and known unoccupied space, in adjacent map-points. When a frontier point is found a new BFS is performed. Starting at the current point it extracts all adjacent frontier points and connects them to one frontier. It is saved in the Frontier-Open-List. The benefit of using WFD is that by maintaining these lists, the algorithm only needs to incorporate points that have not been iterated yet. In fact, the algorithm keeps track of detected open space and frontiers using the above mentioned list. Following iterations make use of this information because they do not need to consider previously known regions any more.

Exploration does not only consist of finding frontiers. Based on the frontiers a target location for the robot needs to be selected. Navigating to the closest frontier points is the simplest frontier-based exploration strategy. For this project, however, a more sophisticated implementation is chosen. Instead of sending the robot to the closest frontier point, a small search is performed to find suitable destinations. Generally, a small region around each frontier cell is searched. The goal is to give each frontier point a rating on how well it is suited as a destination. For this purposes an evaluation function has been implemented. It incorporates three features that are used to evaluate goals to be sent to the navigation stack. They are :
\begin{enumerate}
\item{Number of obstacles in the area}
\item{Distance to the previous goal}
\item{Distance between frontier and robot location}
\end{enumerate}
The first two features have a weight one, whereas the third is assigned a weight of five. This configuration turned out to create the best results (see Subsection \ref{sec:navEval}) The weighted sum of the features is used to find the goal to be sent to the navigation stack. This evaluation function enables the algorithm to select a reachable destination for the robot in almost every case. The frontier cells are sent to the move\_base in the order of the rating. When the navigation stack fails to steer the robot to a designated goal within a time constraint, the next best goal replaces the previous. The information gain is still next to optimal since the selected goal is preferred to be close to frontier points. At last the second feature makes sure that the robots path is smooth. Selecting a goal near the previous one makes the robot avoid too many loops in the emerging graph.

@@

\subsection{Loop closing effect}
\label{sec:loopClosing}
One of the core properties of the Graph SLAM algorithm is the correction of the pose graph and the map whenever a loop closing is detected. This means that when the robot has been exploring an unknown area and then re-enters an explored area, the transformation between the current pose and the past, re-detected, pose gives essential information about the complete pose graph. It corrects the current belief about environment and position.

@@

\subsection{Future research}
As mentioned in Section \ref{disc:expl} the exploration strategy evolved during the process of testing. This does not mean that it cannot be improved upon. More features like, for instance, loop closings can be taken into account. This would lead to more constraints to be added to the graph. More constraints could then result in an even more accurate assessment of the robot trajectory. This new feature would require to be tuned well, since more constraints also lead to higher space and time complexities. 
\\

The Graph SLAM algorithm proposed by \citep{Grisetti} can be improved as well. Considering the suggested change in the exploration algorithm, the method would need to cope with a larger number of edges and nodes.
The current implementation uses sparse Cholesky factorization to find the most optimal spatial configuration of the trajectory. This method is inherited in the ``Eigen" library \ref{sec:impl}. Although sparse Cholesky factorization solves the linear system in an efficient manner, it is likely that even better algorithms emerge in the future. Since the time complexity of the Graph SLAM method depends almost entirely on the technique used, new methods would directly improve upon the running time of the algorithm.

>>>>

\subsubsection{Stage and RViz}
During the development process of a system running on a robot one crucial part is testing. A ROS stack called Stage enables developers to test code on a simulated robot allowing it to move around in virtual environments. Stage visualizes the robot, its movement and measurements. The user can load individually created maps into Stage and specify the robot's configuration. A second important stack needed especially for validation is RViz. It visualizes all requested data. In Figure \ref{fig:stage_and_rviz} Stage and RViz are presented next to each other. While Stage (Left) shows the real map, RViz (Right) presents the map data as built by the robot.

\subsection{TurtleBot}
\label{subsec:turtle}
The product of this project is designed to run on a mobile robot called TurtleBot (Figure \ref{fig:turtlebot}) \citep{turtlebot}. It is a low-cost, customizable, personal robot. The base of the TurtleBot is an iRobot Create (A). It holds a battery pack and a 150 $degrees/second$ Single Axis Gyro. The model shown in Figure \ref{fig:turtlebot} uses a Microsoft Kinect sensor (B), the one used in this paper uses a laser range sensor, and a laptop (C) to run all the processes. The robot can also be customized using the mounting hardware (D). The robot runs on an open-source SDK based on ROS which integrates all the software needed to get the TurtleBot running. Some capabilities are included in ROS and can be used on the TurtleBot after minor configurations. These include mapping and navigation, for example. 
\begin{figure}[h]
	\centering
		\includegraphics[width=0.50\textwidth]{figures/turtlebot.png}
	\caption{A TurtleBot \citep{turtlebot}}
	\label{fig:turtlebot}
\end{figure}

\subsection{Exploration}
\label{subsec:sotaExplore}
Usually robots localize themselves and navigate through environments using a map which is provided by an external source. In this case navigating from one location to another is an easy task. If the robot does not have access to a map, navigating through an environment is problematic. The problem treated in this paper (Subsection \ref{sec:problem}) assumes an unknown environment and, hence, no map. The robot needs to be able to explore its environment without any knowledge about it. It needs to rely on the incoming laser and odometry data only. In addition to the creation of the map and the localization of itself (the SLAM problem) the robot needs to explore the environment as efficient as possible. A good exploration strategy is defined as one that generates an accurate map in the shortest amount of time.

Ideally the robot updates its internal map of the environment constantly while moving. One idea to efficiently explore the world is to move to locations which maximize the robot's information gain. This target location needs to be extracted from the limited information the robot has collected with its sensors.
To achieve this the robot makes use of the idea of frontier-based exploration, first introduced by Yamauchi \cite{Yamauchi} in 1998. Frontiers are defined to be the boundaries between free and unexplored space. Intuitively, when a robot moves to a frontier it is likely that it takes measurements of previously unexplored space. Given that, new data can be added to its map. The explored region expands once the robot reaches a frontier. Moreover an extended map results in new frontiers. By constantly moving to frontiers the robot expands its knowledge about the environment.

@@

SLAM is a problem because localization requires a map and mapping requires the location of the robot that wants to create a map. There are quite a few approaches that successfully solve the SLAM problem. 
In general there are two interpretations it. The first estimates the current robot pose along with the map. The core of such a system is a filtering technique. New measurements augment and refine the estimate as they become available. Due to their incremental nature these filtering approaches are commonly referred to as on-line SLAM methods. FastSLAM \citep{Montemerlo02} is an example of an on-line SLAM implementation. It uses a Particle Filter to approximate the robot's pose and the landmark positions on the map. Other methods have the goal to find the trajectory of the robot and the map. These solve the full SLAM problem and typically rely on least-square error minimization techniques \cite{Leastsquares} to find adequate estimates.  
An example of such a an approach is Graph SLAM.

\subsubsection{Graph SLAM}
An intuitive way to formulate the SLAM problem is to view it as a pose graph optimization problem \citep{Grisetti}. Each pose along the trajectory of the robot is represented by a node. Edges are labeled with relative spatial information between the connecting nodes. This graph is built incrementally. While the robot is in motion, new nodes are added to the graph. Building up the graph, as explained above, is a straight forward procedure. The essence of the Graph SLAM method, however, is the optimization step. In this step the algorithm tries to find a spatial configuration of the trajectory which is maximally consistent with the system of constraints obtained from the edges. This problem is commonly solved using least-square error minimization techniques \citep{Leastsquares}.

@@

The implemented packages make use of the external library ``Eigen" \citep{eigen}. It provides extensive linear algebra operations. Included are simple matrix operations as well as advanced techniques. Two techniques which are used in this project are Singular Value Decomposition and sparse Cholesky factorization.

\subsection{Graph SLAM}
\label{sec:implSLAM}
The implemented Graph SLAM algorithm is based on Grisetti et al.\citep{Grisetti}. The graph consists of nodes representing poses along the robot trajectory. Odometry and sensor measurements are incorporated into the constraints contained in the edges. Similar to the Graph SLAM procedure explained in Section \ref{subsec:sotaSlam}, an edge connects two adjacent poses of the robot. 
\begin{figure}[h]
	\centering
		\includegraphics[width=0.50\textwidth]{figures/graph.png}
	\caption{An example of a simple graph \citep{Grisetti}}
	\label{fig:turtlebotexample_graph}
\end{figure}
When the graph is created the pose graph is optimized according to the constraints. This ``optimized" trajectory then determines the positions of obstacles on the map. The laser scan data is refined according to the poses they were obtained at. Each edge, thus, is labeled with the spatial difference between two poses \(x_i\) and \(x_j\). Let this error be denoted by \(e_{ij}(x_i, x_j)\). This information is obtained from a combination of odometry data and laser range sensor data. In fact, the raw odometry data is corrected using a laser scan matcher (see Subsection \ref{sec:scan}). 
Figure \ref{fig:turtlebotexample_graph} shows a sketch of a graph created by the algorithm. Two poses \(x_i\) and \(x_j\) are connected by an edge (Solid arrow). The ellipse at the top illustrates the label of the edge.

This Graph SLAM approach calculates the most likely path as follows:
First it finds the Jacobians \(A\) and \(B\) of every two estimated poses connected by an edge.
\begin{equation}
\label{eq:jacobi_a}
	A_{ij} = \frac{\partial e_{ij}(x)}{\partial x_i}
\end{equation}
\begin{equation}
\label{eq:jacobi_b}
	B_{ij} = \frac{\partial e_{ij}(x)}{\partial x_j}
\end{equation}
Note that in Equations \ref{eq:jacobi_a} and \ref{eq:jacobi_b} the x represents the current guess of the concerning robot pose. As explained above, \(e_{ij}(x)\) is the error between the two poses \(x_i\) and \(x_j\). The Jacobians \(A\) and \(B\) are then added to a matrix \(H\) (Equation \ref{eq:H}).
\begin{equation}
\label{eq:H}
\begin{aligned}
	H_{[ii]} += A^{T}_{ij} \Omega_{ij} A_{ij} \\
	H_{[ij]} += A^{T}_{ij} \Omega_{ij} B_{ij} \\
	H_{[ji]} += B^{T}_{ij} \Omega_{ij} A_{ij} \\
	H_{[jj]} += B^{T}_{ij} \Omega_{ij} B_{ij} \\
\end{aligned}
\end{equation}
In this equation $\Omega_{ij}$ is the information matrix of a measurement between \(x_i\) and \(x_j\). It is defined as shown in Equation \ref{eq:omega}.
Sigma is the $3 \times 3$ covariance matrix of the measurements. In our implementation the diagonal of $\Sigma$ is filled with arbitrary values.
\begin{equation}
\label{eq:omega}
\Omega = \Sigma^{-1}
\end{equation}
In addition to the matrix $H$ a coefficient vector \(b\) is kept up to date as shown in Equation \ref{eq:b}. This vector together with the \(H\) form a set of linear constraints.
\begin{equation}
\label{eq:b}
\begin{aligned}
	b_{[i]} += A^{T}_{ij} \Omega_{ij} e_{ij} \\
	b_{[j]} += B^{T}_{ij} \Omega_{ij} e_{ij} 
\end{aligned}
\end{equation}
The optimization of the vector \(x = (x_1, ..., x_n)^T \) according to the constraints specified in \(H\) and \(b\) is the core of this approach. This step is performed using sparse Cholesky factorization.

A new node is added to the graph once the robot moves 0.5 meters. The algorithm optimizes the trajectory like explained once a new node is added.\\
In one special case the algorithm adds an extra edge, connecting to non-adjacent nodes, to the graph. This happens when the distance between the two connected robot poses is smaller than a margin. In fact, this special case enables the algorithm to create loops in the graph and more accurately estimate the robot's trajectory and the map. This procedure is referred to as ``loop closing" (see Subsection \ref{sec:loopClosing}). 
\subsection{Scan matching}
\label{sec:scan}
The scan matcher, used by this program, is necessary to acquire an optimized movement vector between two robot poses. This is needed by the Graph SLAM algorithm (see Subsection \ref{sec:implSLAM}). The scan-matcher realized for this project can be divided into two steps. At the beginning, the two scans are matched making use of raw odometry data, retrieved by the robot's motion sensor. This is improved upon by running an iteration of the iterative closest point algorithm (ICP) \citep{scanmatch}. It tries to align the two scans minimizing the total distance of each point in the first scan to its nearest neighbor in the second.

The general procedure of the algorithm \citep{scanmatch} is to first find the nearest neighboring point in the second scan for all point in the first. To achieve this nearest neighbor search is used.
In this implementation multiple points can have the same corresponding neighbor-point. If there is no point from the other scan within a specified radius of one point, it is excluded. These outliers would influence further calculations negatively.
The next step in the process is to calculate the centroids of both clouds. These are subtracted from every point in the concerning cloud. In that way the clouds are centered at the origin. The coordinates to the calculated points are then added to two separate $2 \times n$ matrices, where $n$ is the number of neighbor pairs. Next the rotation matrix which transforms one cloud to the other is needed. The Singular Value Decomposition (SVD) of the product of the two previously mentioned matrices returns three new matrices $U$, $\Sigma$ and $V$. The required rotation matrix R can be obtained as described in Equation \ref{eq:RVUT}.
\begin{equation}
\label{eq:RVUT}
 R = V * U^T
\end{equation}
The translation vector $t$ is the difference between the two centroids. $R$ and $t$ together form a rigid transformation. This transformation matches the two scans. Moreover it is used to correct the robot's odometry data.

@@

\subsection{Mapping}
\label{sec:mapping}
The map is internally represented as an occupancy grid. It consists of cells which hold information about obstacles, free and unknown space.

The occupancy grid is a structure that is represented by a one-dimensional array. The size of an occupancy grid has to be tuned problem specific. A grid, which is too large can lead to performance problems. If the grid is too small, however, the map will be inaccurate. Fitting a map on a smaller grid results in a low resolution.
The program treats the grid as a two dimensional array. This makes navigating to either side more intuitive. Each grid cell contains an integer value. The value-range for ROS' occupancy grid lies between $-1$ and $100$ \cite{occupancy}. The values indicate the state as follows:
\begin{description}
\item{$-1$} : Unknown space, not yet covered by robot's perception sensors
\item{$0$} : Known free space, robot can drive on it safely
\item{$1-100$} : Probability of being an obstacle, $1$ being a low probability and $100$ being a definite obstacle.
\end{description}
\begin{figure}[htbp]
	\centering
		\includegraphics[width=0.45\textwidth]{figures/Occup.png}
	\caption{A visualized occupancy grid}
	\label{fig:Occupancy}
\end{figure}
Figure \ref{fig:Occupancy} shows an example occupancy grid. The robot stands in the center of it and makes one measurement consisting of several laser scans. Each scan returns the range it was able to reach. The returned range is either the maximally possible range of the laser or a lower value, indicating an obstacle (value of 100). The range is transformed into global coordinates with respect to the robot's location, orientation and the bearing of the laser beam. The resulting point is then transferred into the occupancy grid as an obstacle if the maximum range was not reached. The cells lying in a straight line between the transformed range and the robot are considered free space (value of 0), since the laser was able to pass through this area. All the cells that are not affected by any of the previous steps remain unexplored (value of -1). Naturally, this is the case for cells behind obstacles, out of range and directly behind the robot.

@@

\subsection{Exploration}
\label{sec:wfd}
One part of the assignment is autonomous exploration. The robot has to drive autonomously through an unknown world. This feature is implemented using Wavefront Frontier Detection (WFD \citep{Keidar}).

WFD is a frontier-based exploration algorithm. It uses four different data structures. These store map and frontier information during the search process. They are:
\begin{enumerate}
\item{Map-Open-List}
\item{Map-Close-List}
\item{Frontier-Open-List}
\item{Frontier-Close-List}
\end{enumerate}
To understand the need of these four structures it is necessary to know that WFD is using a breadth first search (BFS), starting at the current robot position. Moreover frontier detection requires the robot to have an internal representation of the current environment. In this implementation an occupancy grid (Subsection \ref{sec:mapping}) is used.
Upon launching the algorithm it first searches for unoccupied space around the current robot location. These grid cells are saved in the Map-Open-List. 
Frontiers are found by making use of the occupancy grid's structure. When a frontier point is found a new BFS is performed. Starting at the current point it extracts all adjacent frontier points and connects them to one frontier. It is saved in the Frontier-Open-List. The benefit of using WFD is that by maintaining these lists, the algorithm only needs to incorporate points that have not been iterated yet. In fact, the algorithm keeps track of detected open space and frontiers using the above mentioned list. Following iterations make use of this information because they do not need to consider previously known regions any more.

Exploration does not only consist of finding frontiers. Based on the frontiers a target location for the robot needs to be selected. Navigating to the closest frontier points is the simplest frontier-based exploration strategy. For this project, however, a more sophisticated implementation is chosen. Instead of sending the robot to the closest frontier point, a small search is performed to find suitable destinations. Generally, a small region around each frontier cell is searched. The goal is to give each frontier point a rating on how well it is suited as a destination. For this purposes an evaluation function has been implemented. It incorporates three features that are used to evaluate goals to be sent to the navigation stack. They are :
\begin{enumerate}
\item{Number of obstacles in the area}
\item{Distance to the previous goal}
\item{Distance between frontier and robot location}
\end{enumerate}
The first two features have a weight one, whereas the third is assigned a weight of five. This configuration turned out to create the best results (see Subsection \ref{sec:navEval}) The weighted sum of the features is used to find the goal to be sent to the navigation stack. This evaluation function enables the algorithm to select a reachable destination for the robot in almost every case. The frontier cells are sent to the navigation stack in the order of the rating. When the navigation stack fails to steer the robot to a designated goal within a time constraint, the next best goal replaces the previous. The information gain is still next to optimal since the selected goal is preferred to be close to frontier points. At last the second feature makes sure that the robot's path is smooth. Selecting a goal near the previous one makes the robot avoid too many loops in the emerging graph.

@@

\subsection{Loop closing effect}
\label{sec:loopClosing}
One of the core properties of the Graph SLAM algorithm is the correction of the pose graph and the map whenever a loop closing is detected. This means that when the robot is exploring an unknown area and then re-enters an explored area, the transformation between the current pose and the past, re-detected, pose gives essential information about the complete pose graph. It corrects the current belief about environment and position.

@@

\subsection{Future research}
\label{sec:future}
As mentioned in Section \ref{disc:expl} the exploration strategy evolved during the process of testing. This does not mean that it cannot be improved. More features like, for instance, loop closings can be taken into account. This would lead to more constraints to be added to the graph. More constraints could then result in an even more accurate assessment of the robot trajectory. This new feature would require to be tuned well, since more constraints also lead to higher space and time complexities. 
\\

The Graph SLAM algorithm proposed by \citep{Grisetti} can be improved as well. Considering the suggested change in the exploration algorithm, the method would need to cope with a larger number of edges and nodes.
The current implementation uses sparse Cholesky factorization to find the most optimal spatial configuration of the trajectory. This method is inherited in the ``Eigen" library (Subsection \ref{sec:struct}). Although sparse Cholesky factorization solves the linear system in an efficient manner, it is likely that even better algorithms emerge in the future. Since the time complexity of the Graph SLAM method depends almost entirely on the technique used, new methods would directly improve the running time of the algorithm.
